\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-28 tex/paper.tex
save

\begin{figure}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$
             and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.
             Adversarial training was performed for $2000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=100$.}
    \label{fig:toy-training}
\end{figure}

@@

Assuming a uniform prior over the interval ${\cal Z} = [0;4]$, we start by
generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
neural network classifier $f$ minimizing $L_f(\theta_f)$ without adversarial
training. As shown in the upper plots of Figure~\ref{fig:toy}, the resulting
classifier is not pivotal, as the conditional probability densities of its
decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
the nuisance. While not shown here, a classifier trained only from data
generated at the nominal value $Z=0$ would also not be pivotal.

By contrast, when using Algorithm~\ref{alg:adversarial-training} for training
$f$ jointly with a mixture density network $r$ modeling the conditional
distribution of the nuisance as observed through $f$, the resulting classifier
becomes effectively almost independent on $Z$, as shown in the lower plots of
Figure~\ref{fig:toy}. In particular, the conditional probability densities of
the decision scores $f(X;\theta_f)$ are now very similar to each other, indicating only a
small  dependency on the nuisance, as theoretically expected.
\glnote{Discuss Figure~\ref{fig:toy-training}}

>>>>

\begin{figure}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for ${\cal L}_f(\theta_f)$, ${\cal L}_r(\theta_f, \theta_r)$
             and ${\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r)$.
             Adversarial training was performed for $2000$ iterations, mini-batches of size $M=128$, $K=1$ and $\lambda=100$.}
    \label{fig:toy-training}
\end{figure}

@@

Assuming a uniform prior over the interval ${\cal Z} = [0;4]$, we start by
generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build a
neural network classifier $f$ minimizing ${\cal L}_f(\theta_f)$ without adversarial
training. As shown in the upper plots of Figure~\ref{fig:toy}, the resulting
classifier is not pivotal, as the conditional probability densities of its
decision scores $f(X;\theta_f)$ show large discrepancies between values $z$ of
the nuisance. While not shown here, a classifier trained only from data
generated at the nominal value $Z=0$ would also not be pivotal.

By contrast, when using Algorithm~\ref{alg:adversarial-training} for training
$f$ jointly with a mixture density network $r$ modeling the conditional
distribution of the nuisance as observed through $f$, the resulting classifier
becomes effectively almost independent on $Z$, as shown in the lower plots of
Figure~\ref{fig:toy}. In particular, the conditional probability densities of
the decision scores $f(X;\theta_f)$ are now very similar to each other,
indicating only a small  dependency on the nuisance, as theoretically expected.
The dynamics of adversarial training is illustrated in
Figure~\ref{fig:toy-training}, where the losses ${\cal L}_f$, ${\cal L}_r$ and
${\cal L}_f - \lambda {\cal L}_r$ are evaluated after each iteration of
Algorithm~\ref{alg:adversarial-training}. In the first $200$ iterations, we
observe that $f$ is worse than random because it rather tries to maximizes the
adversarial loss ${\cal L}_r$ by producing decision scores $f(X;\theta_f)$ that
are very unlikely under the current model $p_{\theta_r}$, at the expense of
making bad predictions with respect to the classification problem. As learning
goes and $p_{\theta_r}$ becomes more accurate, minimizing $E$ requires making
predictions that are more accurate, hence decreasing ${\cal L}_f$, or that are
less dependent on $Z$, hence shaping $p_{\theta_r}$ so that it becomes more
uniform. Indeed, ${\cal L}_f$ eventually start decreasing, while remaining lower
bounded by $\min_{\theta_f} {\cal L}_f(\theta_f)$ as approximated by the dashed
line in the first plot. Similarly,  ${\cal L}_r$ increasingly tends towards the
differential entropy $H(p_Z) = \log(4)$ of the uniform distribution over ${\cal
Z}$, as shown by the dashed line in the second plot.
